Convergence rate of MAP estimates \\ for the exponential family

### 2021-11-09 aistats.tex
Clean beginning of App.A with feedback from Simon/Damien

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
As mentioned in the main text, the centered gaussian $\cN(0,\sigma^2)$ has sufficient statistic $T(X)=X^2$ which follows a gamma distribution $\Gamma\paren{\half, \inv{2 \sigma^2}}$. 
MLE and MAP only depend on $T(X)$, not $X$, so their performance only distributions of $T(X)$. 
In this section we derive results for samples  from a general gamma distribution $X \sim \Gamma(\alpha,\beta)$ with known shape parameter $\alpha$, but unknown rate parameter $\beta$.
Results for the Gaussian follow by taking $\alpha=\half$,
 and we also immediately have results for exponential distributions by taking $\alpha=1$.

\begin{theorem}[MLE Upper Bound]
	Consider an exponential family such that $T(X)$ is a gamma $\Gamma(\alpha, \beta)$ with known shape $\alpha$.
	the expected KL between $\mu_*$ and the MLE $\hat\mu_n$	is infinite when $\alpha n\leq 1$ and otherwise upper bounded by
	\begin{align}
		 \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} + \frac{1}{n(n \alpha-1)}  \; .
	\end{align}
\end{theorem}

This theorem applies to Gaussian variance since 
\TODO{explain how this corresponds to theorem 4.1}

\subsection{Gamma distribution}
The density of $\Gamma(\alpha,\beta)$ reads
\alignn{
	p(x) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x} \; .
}
When $\alpha$ is known, it can be cast as an exponential family~\eqref{eq:def_expfamily} with sufficient statistic $T(x)=x$, domain $\cX = \real_+$ and base measure $\nu(x)\propto x^{\alpha-1}$. Then the natural parameter is $\nat = -\beta < 0$ and the log-partition function is
\alignn{
	\logpart (\nat) = -\alpha \log(-\nat)  + \log\Gamma(\alpha) \; .
}
From there we find that the mean parameter is $\mu = \frac{\alpha}{-\nat} >0$ and the entropy has the same form as the log-partition $\conj(\mu) = - \alpha \log(\mu) + \cst$.
This means that the primal and dual Bregman divergences have the same form as well
\begin{align}
	\bregmanconj(\m_*; \m_n) 
	&= \alpha \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \alpha \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \alpha \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \alpha \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
\begin{wrapfigure}[14]{r}{0.4\textwidth}
	\centering
	\includegraphics[width=.38\textwidth]{phi.pdf}
	\caption{%
	Illustration of $\phi(z)$ (orange) and its upper bound $\phi(z) + \phi\paren{z^{-1}}$ (grey).
	They both are barriers near $0$.
	}
	\label{fig:phi}
\end{wrapfigure}
where these 2 lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := z - 1 - \log(z),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To derive the upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n )
	&:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) \nonumber\\
	&=  \alpha \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman} \\
	&= \alpha\phi\paren{\frac{\mu_*}{\mu_n}} + \alpha\phi\paren{\frac{\mu_n}{\mu_*}}
}
Writing $z=\frac{\mu_*}{\mu_n}$ this is equivalent to
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{z^{-1}}
	&=  z -1  + \inv{z} -1
	= \frac{(z-1)^2}{z} \; ,
}
and as illustrated by the grey upper bound in Figure~\ref{fig:phi}.

\subsection{Proof for the MLE}

	\begin{proof}
	For gamma distributions, since $T(X)=X$ the MLE is a scaled sum of gammas $\hat \mu_n = \inv{n} \sum_i X_i$.
	As such it is also a Gamma with parameter $\Gamma(n \alpha, n \beta)$ and expectation $\frac{n\alpha}{n\beta} = \frac{\alpha}{\beta} = \mu_*$.
	If we consider the ratio $\frac{\MAPm}{\mu_*}$, it is also a Gamma with parameter $\Gamma(n \alpha, n \alpha)$ 
	Its inverse follows an inverse gamma distribution with expectation
	\begin{equation}
		\expect{\frac{\mu_*}{\hat \mu_n}} 
		= \begin{cases}
			\frac{n\alpha}{n \alpha -1 }\ \text{ if } n\alpha > 1, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{equation}
	which implies that for $n\alpha>1$, 
	\begin{align}
		\expect{\frac{\mu_*}{\hat \mu_n}} -1  
		= \frac{n \alpha}{n\alpha -1 } - 1
		= \frac{1}{n \alpha -1}
	\end{align}
	There is also a closed form solution for the expected logarithm of a Gamma.
	Indeed a the sufficient statistic of a Gamma is $(X, \log(X))$, so one can apply formula~\eqref{eq:mirror-map} on the log-partition of a Gamma to get
	\begin{align}
		\expect{\log \frac{\hat \mu_n}{\mu_*}} 
		= \psi\paren{n \alpha} - \log\paren{n \alpha},
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	%Closed forms also exist for non-central chi-squares, see for example the note of \citep{pav2015moments}.
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \hat \mu_n) }
	&= \alpha \expect{\frac{\mu_*}{ \hat \mu_n} - 1 + \log \left(\frac{\hat \mu_n}{\mu_*} \right) },
	\\
	& =
	\begin{cases}
		\alpha \paren{ \frac{1}{n\alpha - 1} + \psi\paren{n \alpha} - \log\paren{n \alpha} }, \ \text{ if } n\alpha>1, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	Surprisingly, for a gaussian variance where $\alpha = \half$, we need $3$ samples or more for the expected loss to be bounded.
	When the expectation is finite, we can get a more interpretable formula using known \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function},
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x} = - \inv{x} + \inv{2x}	\; ,
	\end{align}
	giving, for $n \alpha >1 $,
	\begin{align}
		\frac{\alpha}{n\alpha-1} - \frac{\alpha}{n\alpha}
		&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
		\leq \frac{\alpha}{n\alpha-1} - \frac{\alpha}{n\alpha} + \frac{\alpha}{2n\alpha}
		\\
		\iff
			\frac{1}{n(n \alpha-1)}
			&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} + \frac{1}{n(n \alpha-1)} \; ,
	\end{align}
	so we get a $\Omega(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	\end{proof}

\subsection{Multivariate MLE}

@@

\subsection{Bounding the expected natural parameters for the MAP}
Before proving a convergence rate for the MAP, we need to bound the expectation of its inverse, hence the following lemma.
We introduce the notation $(z)_+ = \max(0,z)$.
\begin{lemma}[Expected MAP natural parameter]\label{lem:expected-map-natural-parameter-gaussian}
	Define the variable $a = n_0 \frac{\mu_0}{\mu^*}$, which characterizes 
	the importance of the prior relative to the true parameters.
	The expectation of the natural parameter of a MAP of $\Gamma(\alpha,\beta)$ is bounded by
\begin{equation}
		\frac{n_0+n}{a+n}
		=\frac{\mu^*}{\mu_n}
		\leq \expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		\leq
		\frac{n_0 +n}{a+(n-\inv{\alpha} )_+} \; , \forall n\geq 0 \; .
		\label{eq:lemma_nat_bound}
	\end{equation}
\end{lemma}

\begin{proof}
	The lower bound can be readily obtained by applying Jensen's inequality 
	to the convex function $x\mapsto\inv{x}$ for $x>0$.
	The upper bound requires more work.
	To start, let us plug in the definition of $\MAPm$
\alignn{
		\expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		= \expect{\frac{(n_0+n) \mu^*}{n_0 \mu_0 + \sum_i X_i}}  
		= \expect{\frac{n_0+n}{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i}{ \mu^*} }} 
		= \expect{\frac{n_0+n}{a + \Gamma(n\alpha, \alpha)}} \; ,
}
	where  $\sum_i \frac{X_i^2}{ \mu^*} \sim \Gamma(n\alpha, \alpha)$ is a Gamma random variable and $a=n_0 \frac{\mu_0}{\mu^*}$. 
	Further note that 
\alignn{
	\expect{\frac{n_0+n}{a + \Gamma(n\alpha, \alpha)}} 
	= \frac{n_0 + n}{a} \expect{\frac{1}{1 + \Gamma(n \alpha , a \alpha) }} \; .
}
This kind of integrals can be expressed with {generalized exponential integral functions}
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; ,
	\end{align}
with \href{http://dlmf.nist.gov/8.19.E4}{the formula} \citep[Eq.~8.19.4]{DLMF}
\alignn{
	\expect{\frac{1}{1 + \Gamma(\alpha , \beta) }}
	= \frac{\beta^{\alpha}}{\Gamma(\alpha)} \int_0^\infty \frac{x^{\alpha-1}e^{-\beta x} }{1+x}  dx 
	= \beta e^{\beta} E_\alpha (\beta) \; .
	\label{eq:expect1+gamma}
}
Overall we get
\alignn{
\expect{\frac{\mu^*}{\MAPm}}  = (n_0 +n) \alpha e^{a \alpha} E_{n\alpha} ( a \alpha) 
}
	Now our goal is to bound this generalized exponential integral with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	For instance , we have \href{https://dlmf.nist.gov/8.19.E21}{the general bound} \citep[Eq.~8.19.21]{DLMF}
	\begin{align}
		e^x E_k(x) 
		&\leq \inv{x + k - 1} , \ \forall k>1\\
		\iff \expect{\frac{\mu^*}{\MAPm}} 
		&\leq \frac{(n_0 +n)\alpha}{a\alpha + n\alpha - 1}
		= \frac{n_0 +n}{a + n - \inv{\alpha}} , \ \forall n >\inv{\alpha} \; .
	\end{align}
We are left with a special case when  $n\leq \alpha$. 
Then we can use the trivial bound 
\begin{align}
	\expect{\inv{a + X^2}} < \inv{a}.
	\label{eq:trivial_bound}
\end{align}
to conclude the proof.
\end{proof}

@@

\subsection{Proof of MAP Bound}

@@

There are several ways to write down the convergence rate. The Gaussian variance example can be written in a particularly simple form, so we give  it a special treatment in \S\ref{app:proof-MAP-gaussian}, corresponding to the theorem displayed in the main text.
We make a more general statement about Gamma distributions in \S\ref{app:proof-MAP-gamma}.

@@

\subsubsection{Gamma with known shape}
\label{app:proof-MAP-gamma}

\begin{theorem}[MAP Bound]
Consider an exponential distribution with sufficient statistics coming from a gamma distribution $\Gamma(\alpha, \beta^*)$ with mean parameter $\mu^* = \frac{\alpha}{\beta^*}$.
The expected symmetrized Bregman~\eqref{eq:symmetrized_bregman} of the MAP  with prior hyper-parameters $(n_0,\mu_0)$ is upper bounded as
\alignn{
    \forall n\geq \inv{\alpha},\quad \quad
	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq\frac{1}{n_0+n} + 
	\frac{\alpha\paren{\frac{\mu_0}{\mu^*} - \inv{\alpha n_0} - 1 }^2 }{\paren{1+\frac{n}{n_0}}\paren{\frac{\mu_0}{\mu^*} +  \frac{n - \inv{\alpha}}{n_0}}}
}
\end{theorem} 
Note that the second term vanishes when $\mu_0 = \mu^* \paren{1 + \inv{\alpha n_0}}$.
Expressions for $n\alpha < 1$ are less elegant so we do not write them down here. 
\TODO{try simplification with wolfram alpha}.

>>>>

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
As mentioned in the main text, the centered gaussian $\cN(0,\sigma^2)$ has sufficient statistic $T(X)=X^2$ which follows a gamma distribution $\Gamma\paren{\half, \inv{2 \sigma^2}}$.

In general, if $X$ is part of the exponential family,  $T(X)$ is part of the natural exponential family with the appropriate support and base measure, and with the same log-partition function as $X$, up to constants.
MLE and MAP only depend on $T(X)$, not $X$, so their performance only depends on the distributions of $T(X)$.

In this section we derive results for samples  from a general gamma distribution $X \sim \Gamma(\alpha,\beta)$ with known shape parameter $\alpha$, but unknown rate parameter $\beta$.
Results for the Gaussian follow by taking $\alpha=\half$,
We also immediately get results for exponential distributions by taking $\alpha=1$.
For instance for the MLE we derive the following theorem:

\begin{theorem}[MLE Upper Bound]
	\label{thm:gammaMLE}
	Consider an exponential family such that $T(X)$ is a gamma $\Gamma(\alpha, \beta)$ with known shape $\alpha$.
	the expected KL between $\mu_*$ and the MLE $\hat\mu_n$	is infinite when $\alpha n\leq 1$ and otherwise upper bounded by
	\begin{align}
		 \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} + \frac{1}{n(n \alpha-1)}  \; .
	\end{align}
\end{theorem}

To obtain the result for Gaussian variance, theorem~\ref{thm:varianceMLE}, simply plug-in $\alpha=\half$.

In this section, we review useful properties of the gamma distribution and associated Bregman divergence in \S\ref{app:gamma}. 
Then we prove theorem~\ref{thm:gammaMLE} in \S\ref{app:gammaMLE}.
Then we prove an extension of theorem~\ref{thm:varianceMLE} in \S\ref{app:multivariateMLE}.
Then we prove a useful lemma about the expectation of the natural parameter of the MAP in \S\ref{app:nat-bound}.
Finally we prove upper bounds for the MAP in \S\ref{app:MAP-bound}.

\subsection{Gamma Distribution}
\label{app:gamma}
The density of $\Gamma(\alpha,\beta)$ reads
\alignn{
	p(x) = \frac{\beta^\alpha}{\Gamma(\alpha)} x^{\alpha-1} e^{-\beta x} \; .
}
When $\alpha$ is known, it can be cast as an exponential family~\eqref{eq:def_expfamily} with sufficient statistic $T(x)=x$, domain $\cX = \real_+$ and base measure $\nu(x)\propto x^{\alpha-1}$. Then the natural parameter is $\nat = -\beta < 0$ and the log-partition function is
\alignn{
	\logpart (\nat) = -\alpha \log(-\nat)  + \log\Gamma(\alpha) \; .
}
From there we find that the mean parameter is $\mu = \frac{\alpha}{-\nat} >0$ and the entropy has the same form as the log-partition $\conj(\mu) = - \alpha \log(\mu) + \cst$.
This means that the primal and dual Bregman divergences have the same form as well
\begin{align}
	\bregmanconj(\m_*; \m_n) 
	&= \alpha \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \alpha \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \alpha \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \alpha \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
\begin{wrapfigure}[14]{r}{0.4\textwidth}
	\centering
	\includegraphics[width=.38\textwidth]{phi.pdf}
	\caption{%
	Illustration of $\phi(z)$ (orange) and its upper bound $\phi(z) + \phi\paren{z^{-1}}$ (grey).
	They both are barriers near $0$.
	}
	\label{fig:phi}
\end{wrapfigure}
where these 2 lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := z - 1 - \log(z),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To derive the upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n )
	&:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) \nonumber\\
	&=  \alpha \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman} \\
	&= \alpha\phi\paren{\frac{\mu_*}{\mu_n}} + \alpha\phi\paren{\frac{\mu_n}{\mu_*}}
}
Writing $z=\frac{\mu_*}{\mu_n}$ this is equivalent to
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{z^{-1}}
	&=  z -1  + \inv{z} -1
	= \frac{(z-1)^2}{z} \; ,
}
and as illustrated by the grey upper bound in Figure~\ref{fig:phi}.

\subsection{Proof for the MLE}
	\label{app:gammaMLE}

	\begin{proof}
	Since $T(X)$ follows a gamma distribution $\Gamma(\alpha, \beta)$, the MLE is a scaled sum of gammas $\hat \mu_n = \inv{n} \sum_i T(X_i)$.
	As such it is also a gamma with parameter $\Gamma(n \alpha, n \beta)$ and expectation $\frac{n\alpha}{n\beta} = \frac{\alpha}{\beta} = \mu_*$.
	If we consider the ratio $\frac{\MAPm}{\mu_*}$, it is also a gamma with parameter $\Gamma(n \alpha, n \alpha)$.
	Its inverse follows an inverse gamma distribution with expectation
	\begin{equation}
		\expect{\frac{\mu_*}{\hat \mu_n}} 
		= \begin{cases}
			\frac{n\alpha}{n \alpha -1 }\ \text{ if } n\alpha > 1, \\
			+\infty \  \text{ otherwise. }
		\end{cases}
	\end{equation}
	which implies that for $n\alpha>1$, 
	\begin{align}
		\expect{\frac{\mu_*}{\hat \mu_n}} -1  
		= \frac{n \alpha}{n\alpha -1 } - 1
		= \frac{1}{n \alpha -1}
	\end{align}
	There is also a closed form solution for the expected logarithm of a gamma.
	Indeed a the sufficient statistic of a gamma is $(X, \log(X))$, so one can apply formula~\eqref{eq:mirror-map} on the log-partition of a gamma to get
	\begin{align}
		\expect{\log \frac{\hat \mu_n}{\mu_*}} 
		= \psi\paren{n \alpha} - \log\paren{n \alpha},
	\end{align}
	where $\psi$ is the \href{https://en.wikipedia.org/wiki/Digamma_function}{digamma function.}
	%Closed forms also exist for non-central chi-squares, see for example the note of \citep{pav2015moments}.
	Consequently the suboptimality of the MLE has a closed form solution
	\begin{align}
	\expect{\bregmanconj( \mu_*; \hat \mu_n) }
	&= \alpha \expect{\frac{\mu_*}{ \hat \mu_n} - 1 + \log \left(\frac{\hat \mu_n}{\mu_*} \right) },
	\\
	& =
	\begin{cases}
		\alpha \paren{ \frac{1}{n\alpha - 1} + \psi\paren{n \alpha} - \log\paren{n \alpha} }, \ \text{ if } n\alpha>1, \\
			+\infty \  \text{ otherwise. }
	\end{cases}
	\end{align}
	Surprisingly, for a gaussian variance where $\alpha = \half$, we need $3$ samples or more for the expected loss to be bounded.
	When the expectation is finite, we can get a more interpretable formula using known \href{https://en.wikipedia.org/wiki/Digamma_function#Inequalities}{bounds on the digamma function},
	\begin{align}
		-\inv{x} \leq \psi(x) - \log(x) \leq -\inv{2x} = - \inv{x} + \inv{2x}	\; ,
	\end{align}
	giving, for $n \alpha >1 $,
	\begin{align}
		\frac{\alpha}{n\alpha-1} - \frac{\alpha}{n\alpha}
		&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
		\leq \frac{\alpha}{n\alpha-1} - \frac{\alpha}{n\alpha} + \frac{\alpha}{2n\alpha}
		\\
		\iff
			\frac{1}{n(n \alpha-1)}
			&\leq \expect{\bregmanconj( \mu_*; \hat \mu_n) }
			\leq \inv{2n} + \frac{1}{n(n \alpha-1)} \; ,
	\end{align}
	so we get a $\Omega(n^{-2})$ lower bound and a $O(n^{-1}) + O(n^{-2})$ upper bound.
	\end{proof}

\subsection{Multivariate MLE}
\label{app:multivariateMLE}

@@

\subsection{Bounding the expected natural parameters for the MAP}
\label{app:nat-bound}
Before proving a convergence rate for the MAP, we need to bound the expectation of its inverse, hence the following lemma.
We introduce the notation $(z)_+ = \max(0,z)$.
\begin{lemma}[Expected MAP natural parameter]\label{lem:expected-map-natural-parameter-gaussian}
	Define the variable $a = n_0 \frac{\mu_0}{\mu^*}$, which characterizes 
	the importance of the prior relative to the true parameters.
	The expectation of the natural parameter of a MAP of $\Gamma(\alpha,\beta)$ is bounded by
\begin{equation}
		\frac{n_0+n}{a+n}
		=\frac{\mu^*}{\mu_n}
		\leq \expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		\leq
		\frac{n_0 +n}{a+(n-\inv{\alpha} )_+} \; , \forall n\geq 0 \; .
		\label{eq:lemma_nat_bound}
	\end{equation}
\end{lemma}

\begin{proof}
	The lower bound can be readily obtained by applying Jensen's inequality 
	to the convex function $x\mapsto\inv{x}$ for $x>0$.
	The upper bound requires more work.
	To start, let us plug in the definition of $\MAPm$
\alignn{
		\expect{\frac{\mu^*}{\MAPm}} 
		= \expect{\frac{\MAPt}{\nat^*}} 
		= \expect{\frac{(n_0+n) \mu^*}{n_0 \mu_0 + \sum_i X_i}}  
		= \expect{\frac{n_0+n}{n_0 \frac{\mu_0}{\mu^*}+ \sum_i \frac{X_i}{ \mu^*} }} 
		= \expect{\frac{n_0+n}{a + \Gamma(n\alpha, \alpha)}} \; ,
}
	where  $\sum_i \frac{X_i^2}{ \mu^*} \sim \Gamma(n\alpha, \alpha)$ is a gamma random variable and $a=n_0 \frac{\mu_0}{\mu^*}$. 
	Further note that 
\alignn{
	\expect{\frac{n_0+n}{a + \Gamma(n\alpha, \alpha)}} 
	= \frac{n_0 + n}{a} \expect{\frac{1}{1 + \Gamma(n \alpha , a \alpha) }} \; .
}
This kind of integrals can be expressed with {generalized exponential integral functions}
	\begin{align}
		E_k(z) = \int_1^\infty \frac{e^{-z t} }{t^k} dt \; ,
	\end{align}
with \href{http://dlmf.nist.gov/8.19.E4}{the formula} \citep[Eq.~8.19.4]{DLMF}
\alignn{
	\expect{\frac{1}{1 + \Gamma(\alpha , \beta) }}
	= \frac{\beta^{\alpha}}{\Gamma(\alpha)} \int_0^\infty \frac{x^{\alpha-1}e^{-\beta x} }{1+x}  dx 
	= \beta e^{\beta} E_\alpha (\beta) \; .
	\label{eq:expect1+gamma}
}
Overall we get
\alignn{
\expect{\frac{\mu^*}{\MAPm}}  = (n_0 +n) \alpha e^{a \alpha} E_{n\alpha} ( a \alpha) 
}
	Now our goal is to bound this generalized exponential integral with simpler functions.
	Fortunately, mathematicians have been working on these integrals for decades.
	For instance , we have \href{https://dlmf.nist.gov/8.19.E21}{the general bound} \citep[Eq.~8.19.21]{DLMF}
	\begin{align}
		e^x E_k(x) 
		&\leq \inv{x + k - 1} , \ \forall k>1\\
		\iff \expect{\frac{\mu^*}{\MAPm}} 
		&\leq \frac{(n_0 +n)\alpha}{a\alpha + n\alpha - 1}
		= \frac{n_0 +n}{a + n - \inv{\alpha}} , \ \forall n >\inv{\alpha} \; .
	\end{align}
We are left with a special case when  $n\leq \alpha$. 
Then we can use the trivial bound 
\begin{align}
	\expect{\inv{a + X^2}} < \inv{a}.
	\label{eq:trivial_bound}
\end{align}
to conclude the proof.
\end{proof}

@@

\subsection{Proof of MAP Bound}
\label{app:MAP-bound}

@@

There are several ways to write down the convergence rate. The Gaussian variance example can be written in a particularly simple form, so we give  it a special treatment in \S\ref{app:proof-MAP-gaussian}, corresponding to the theorem displayed in the main text.
We make a more general statement about gamma distributions in \S\ref{app:proof-MAP-gamma}.

@@

\subsubsection{Gamma with Known Shape}
\label{app:proof-MAP-gamma}

\begin{theorem}[MAP Bound]
Consider an exponential distribution with sufficient statistics coming from a gamma distribution $\Gamma(\alpha, \beta^*)$ with mean parameter $\mu^* = \frac{\alpha}{\beta^*}$.
The expected symmetrized Bregman~\eqref{eq:symmetrized_bregman} of the MAP  with prior hyper-parameters $(n_0,\mu_0)$ is upper bounded as
\alignn{
    \forall n\geq \inv{\alpha},\quad \quad
	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq\frac{1}{n_0+n} + 
	\frac{\alpha\paren{\frac{\mu_0}{\mu^*} - \inv{\alpha n_0} - 1 }^2 }{\paren{1+\frac{n}{n_0}}\paren{\frac{\mu_0}{\mu^*} +  \frac{n - \inv{\alpha}}{n_0}}}
}
\end{theorem} 
Note that the second term vanishes when $\mu_0 = \mu^* \paren{1 + \inv{\alpha n_0}}$.
Expressions for $n\alpha < 1$ are less elegant. See below \TODO{Damien try this on Matlab?}

\alignn{
 	\expect{\cS_\conj(\mu^*, \hat \mu_n)}
	\leq \frac{n_0 + n}{a + n +\inv{\alpha}} \paren{1 + \inv{a \alpha}} -1 + \frac{a - n_0}{n_0+n}
	\quadtext{where}
	 a:=n_0\frac{\mu_0}{\mu^*} \; .
}
